---
title: "Main use case of aconr"
author: "Andrey Tovchigrechko"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Main use case of aconr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
The main intended use case for the `aconr` package is the creation of nested configuration structures for the
analysis pipelines. `aconr` stands for Active Config in R.

This vignette assumes that you have read the documentation on `acon` class constructor, which is the main entry point of the
`aconr` package. You can access it with `?aconr::acon` from the R session once you install the package.

With the `acon` class constructor, a software pipeline can define
the default instance of a possibly large and deeply nested config, where values deeper in the
config hierarchy are defined as expressions computed from the values of the keys located higher up in the
hierarchy. To tune the config to their needs, the user can derive another instance of
the entire config by selectively modifying values at arbitrary levels of the hierarchy with
a minimal amount of code that only dereferences the keys to be modified. The config is then passed to
the pipeline that uses it to obtain the parameters. Other values in the config which depend on the modified
values will be appropriately recomputed.

Below, we will show how the default config is constructed, and how it is modified by the user for the specific analysis run.
But first, we will create the "analysis pipeline" that consumes the config.

## The "analysis pipeline"

```{r}
bar <- function(param_x,param_z,param_l=NULL) {
  print("Using parameters in functions 'bar':") 
  print(list(param_x=param_x,param_z=param_z,param_l=param_l))
}

foo <- function(param_s,param_f,param_t,param_p) {
  print("Using parameters in functions 'foo':")
  print(list(param_s=param_s,param_p=param_p))
  do.call(bar,param_t)
}
```

## Define the default config

With `acon`, a config looks like a nested named list or a hierarchy of objects in JSON.
The `acon` constructor replaces the use of `list` constructor, thus providing the resulting
datastructure with the special abilities.

In this imaginary example, we assum that all named elements are parameters that will be
used by functions in the analysis workflow that we will define after we build the config.

This is subcomponent (subtree) of the config. Notice that `param_x` is defined as an expression
based on the value of `param_v`, which is defined higher up in the enclosing branch of this config
snippet.

```{r}
library(aconr)
param_z <- acon(param_v=4,param_w=acon(param_x=param_v*2))
```

The word Active in the full name of this datastructure means that wherever `param_x` is accessed, the 
expression will be recomputed by finding the current value of the `param_v` and multiplying it by 2: 
`r param_z$param_w$param_x`. This is similar to
the behavior of the `recursively expanded variables` in the [Make](https://www.gnu.org/software/make/manual/html_node/Flavors.html) 
program. The lookup of the parameters referenced in the expression happens just like the variable lookup
in R through a chain of enclosing environments. In fact, the `acon` is implemented as a subclass of the R `environment`.

Define the entire config datastructure, reusing the existing `param_z` as a node of the configuration tree in one place, 
and defining new unrelated parameter with the same name `param_z` in another place.

```{r}
config_default <- acon(param_s=1:2,
                 param_f=function(x) sprintf('Argument is %s',paste(x,collapse = ",")),
                 param_t=acon(param_x=param_s*2,
                        param_z=acon(param_k=param_x*4)
                 ),
                 param_p = param_z
)
```

Above, we added `param_f` that has a function as its value that can be used, for example, as a call-back by the analysis code.

The values of the config elements can be read (but not changed) using the usual list element notation:

```{r}
config_default$param_t$param_z$param_k
```

## Define the config with the user modifications

Now, the user wants to modify some values in the config before passing the config to the analysis workflow.
The modification has to be done by constructing new `acon` top-level object and passing the original config
object as the **first unnamed** parameter. Other arguments in that call should be names of those config elements
that the user wants to update with new values. If the old value was itself the `acon` object, the new
value can be a call to the `acon` constructor with the named arguments for **only those elements** which the
user wants to change or, in our example, create (see `param_t=acon(param_l=10)` below that creates `param_l`). 
Such an update can be 
called by referencing the parameters down in a chain of `acon` constructor calls to any depth in the configuration tree.

```{r}
config_user <- acon(config_default,
              param_s=1:3,
              param_t=acon(param_l=10))
```

Notice how the `param_k` value changed because we changed the
`param_s` on which the `param_k` indirectly depends.

```{r}
config_user$param_t$param_z$param_k
```

`config_user` is always created as a deep copy of the `config_default`, which stays the same:

```{r}
config_default$param_t$param_z$param_k
```

You can only modify the elements of the config through the constructor pattern above.
An attempt to assign directly to the element like `config_user$param_s <- 1:4` will trigger an error.

You can use the updated `acon` config datastructure directly in your analysis code. However,
typically your downstream analysis code does not need the ability to update the parameters through the "active" 
mechanism. In that case, we recommend converting the updated config to a regular nested named list before using
it. That removes the dependency of your analysis code on the `acon` object and improves performance when accessing
the confif elements in your code.

```{r}
config_user = as.list(config_user)
```

## Call our analysis with the modified config

```{r}
do.call(foo,config_user)
```
